---
title: "Course Project 4"
author: Irra_Balbuena
date: "May 25, 2019"
output: html_document
---

This is my first Markdown document

Setting Directory
```{r directory}
setwd("C:/Users/Irra Balbuena/Documents/balbuena R")
```

Load some data
```{r}
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
summary(NEI)
summary(SCC)
```

Course Project 4 Questions

1. Total emissions from PM2.5 decreased in the United States from 1999 to 2008.
```{r}
totalNEI <- aggregate(Emissions ~ year, NEI, sum)
totalNEI
totalNEI_year <- c("1999", "2002", "2005", "2008")
```
Here's the barplot for question 1
```{r barplot}
barplot(totalNEI$Emissions, names.arg = totalNEI_year, main = expression("Total US "~ PM[2.5]~ "Emissions by Year"), ylab = expression("Total US "~   PM[2.5] ~ "Emissions"), xlab = "Year", col = "steelblue3")
```

2. Total emissions from PM2.5 decreased in the Baltimore City, Maryland from 1999 to 2008.
```{r, echo=TRUE}
baltimore <- subset(NEI, NEI$fips == "24510")
totalBaltimore <- aggregate(Emissions ~ year, baltimore, sum)
totalBaltimore
totalbal_year <- c("1999", "2002", "2005", "2008")
```
Here's the barplot for question 2
```{r}
barplot(totalBaltimore$Emissions, names.arg = totalbal_year, main = expression("Total Baltimore" ~ PM[2.5]~ "Emissions by Year"), xlab = "Year", ylab = expression("Total Baltimore" ~ PM[2.5]~ "Emissions"), col = "steelblue3")
```

3. The four sources indicated by the type(point, nonpoint, onroad, nonroad) variable have seen decreased and increased in emissions from 1999-2008 for Baltimore City

Load package
```{r}
library(ggplot2)
```

```{r}
baltimore <- subset(NEI, NEI$fips == "24510")
baltimoreType <- aggregate(Emissions ~ year + type, baltimore, sum)
baltimoreType
```
Here's the barplot for question 3
```{r}
ggplot(baltimoreType,aes(factor(year),Emissions,fill=type)) +
    geom_bar(stat="identity") +
    facet_grid(.~type) + 
    labs(x="year", y=expression("Total PM"[2.5]*" Emission (Tons)")) +
    labs(title=expression("PM"[2.5]*" Emissions, Baltimore City 1999-2008 by Source Type"))
```

4. Emissions from coal-related sources changed from 1999-2008

Loading package
```{r}
library(ggplot2)
```

```{r}
SCCcoal <- SCC[grepl("coal", SCC$Short.Name, ignore.case = T),]
NEIcoal <- NEI[NEI$SCC %in% SCCcoal$SCC,]
totalCoal <- aggregate(Emissions ~ year + type, NEIcoal, sum)
totalCoal
```
Here's the barplot for question 4
```{r}
ggplot(totalCoal,aes(factor(year),Emissions,fill=type)) +
    geom_bar(stat="identity") +
    facet_grid(.~type) +
    labs(x="year", y=expression("US PM"[2.5]*"Coal Emission")) + 
    labs(title=expression("Total US PM"[2.5]*"Coal Emissions by Type and Year"))
```

5. Emissions form motor vehicle sources changed from 1999-2008 in Baltimore City

Load package
```{r}
library(ggplot2)
```

```{r}
baltimoreMotor <- subset(NEI, NEI$fips == "24510" & NEI$type == "ON-ROAD")
baltimoreMotorAGG <- aggregate(Emissions ~ year, baltimoreMotor, sum)
baltimoreMotorAGG
```
Here's the barplot for question 5
```{r}
ggplot(baltimoreMotorAGG,aes(factor(year),Emissions)) +
    geom_bar(stat="identity") +
    labs(x="year", y=expression("PM"[2.5]*"Motor Vehicle Emissions")) + 
    labs(title=expression("Baltimore PM"[2.5]*"Motor Vehicle Emissions by Year"))
```

6. Comaprison between Emissions from motor vehicle sources in Baltimore City and Los Angeles Country, California

Load package
```{r}
library(ggplot2)
library(plyr)
```

```{r}
baltLosAngelesMotors <- subset(NEI, NEI$fips %in% c("24510","06037") & NEI$type == "ON-ROAD")
baltLosAngelesMotorsAGG <- aggregate(Emissions ~ year + fips, baltLosAngelesMotors, sum)
baltLosAngelesMotorsAGG$fips <- revalue(baltLosAngelesMotorsAGG$fips, c("06037" = "Los Angeles", "24510" = "Baltimore"))
```
Here's the barplot for question 6
```{r}
ggplot(baltLosAngelesMotorsAGG,aes(factor(year),Emissions,fill=fips)) +
    geom_bar(stat="identity") +
    facet_grid(.~baltLosAngelesMotorsAGG$fips) +  ##each panel plot corresponds to a set of values of variable
    labs(x="year", y=expression("PM"[2.5]*"Motor Vehicle Emissions")) + 
    labs(title=expression("Baltimore and Los Angeles PM"[2.5]*"Motor Vehicle Emissions by Year"))
```
